---
title: "Untitled"
author: "Cora Ferguson"
date: "2023-10-19"
output: html_document
---

---
title: "CMF6 - 24 hr Live Dead HDF Encapsulation"
author: "Cora Ferguson"
date: "2023-09-14"
output: html_document
---

##

This is the experiment for the 24 hour solo encapsulation that I performed with human fibroblasts (P4).


Prep environment and load data in

```{r}
# clear the environment 
rm(list = ls())

# import the libraries that we will use
library(ggplot2)
library(dplyr)
library(here)
library(ggfortify)
```


Now import the data for analysis:
```{r}
df <- read.csv(here("Data", "CMF10_live_dead_counts.csv"), stringsAsFactors = TRUE)

# take a look at the structure of the data
glimpse(df)
```


Now let's add a column to calculate the percent viability for the gels:
```{r}
# calculate percent viability
df_clean <- df %>%
  mutate(total_cells = Dead_cells+Live_cells,
    percent_viability = Live_cells/(total_cells)*100)
```

Now let's graph this.  I will group these by gel type (UV vs Vis) and then calculate the mean viability plus or minus the standard error.


```{r}
viability_df <- df_clean %>%
  select(Gel_ID, total_cells, percent_viability) %>%
  group_by(Gel_ID)%>%
  summarise(mean_viability = mean(percent_viability),
            sd_viability = sd(percent_viability),
            se_viability = sd(percent_viability)/sqrt(sum(percent_viability)))
```

Now let's graph this:

```{r}
viability_plt <- viability_df %>%
  ggplot(aes(x = Gel_ID, y = mean_viability))+
  geom_bar(stat = "identity", position = "dodge")+
  geom_errorbar(aes(ymin = mean_viability - sd_viability,
                       ymax = mean_viability + sd_viability), width = 0.1, position = position_dodge(0.9))+
  theme_bw()+
  labs(x = "Gel Type", y= "Mean Percent Viability") +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 100))

viability_plt

```

Let's run a t test on this to assess the statistical relevance: 
```{r}
# run t-test for mass swelling ratio
t.test(percent_viability ~ Gel_ID, data = df_clean)

# check assumptions:
qqnorm(df_clean$percent_viability, main='Normal')
qqline(df_clean$percent_viability)
```

Let's look at viability as the distance from the outside edge of the gel to the inside edge of the gel

```{r}
vi_dis_df <- df_clean %>%
  select(Light, Image_ID, percent_viability) %>%
  filter(Image_ID <=7) %>%
  mutate(gel_depth = Image_ID*50) %>%
  group_by(Light, gel_depth) %>%
  summarise(mean_viability = mean(percent_viability),
            sd_viability = sd(percent_viability))

vi_dis_plt <- vi_dis_df %>%
  ggplot(aes(x = gel_depth, y = mean_viability, fill = Light))+
  geom_bar(stat = "identity", position = "dodge")+
  geom_errorbar(aes(ymin = mean_viability - sd_viability,
                       ymax = mean_viability + sd_viability), width = 10, position = position_dodge(50))+
  theme_bw()+
  labs(x = "Gel Depth (um)", y= "Mean Percent Viability (%)") +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 100))+
  scale_x_continuous(expand = c(0,0), limits = c(-25,400))

vi_dis_plt

```




